\textbf{Machine learning classification using self-reference-based OTU clustering}

### 2023-04-13 submission/manuscript.Rmd
fix typos

1.  OptiFit Self: The preprocess data was split into the training and
    testing sets. The training set was clustered into OTUs using
    OptiClust, then the test set was fit to the OTUs of the training set
    using the OptiFit algorithm @sovacool2022. The OptiFit algorithm
    was run with method open so that any sequences that did not map to
    the existing OTU clusters would form new OTUs. The data was then
    subsampled to 10,000 reads and any novel OTUs from the test set were
    removed. This process was repeated for each of the 100 splits
    resulting in 100 training and testing datasets.
2.  OptiFit Greengenes: Reference sequences from the Greengenes database
    v13_8\_99 @desantis2006 were downloaded and processed with mothur by
    trimming to the V4 region and clustered *de novo* with OptiClust
    @westcott2017. The preprocessed data was fit to the clustered
    reference data using OptiFit with the method open to allow any
    sequences that did not map to the existing reference clusters would
    form new OTUs. The data was then subsampled to 10,000 reads and any
    novel OTUs from the test set were removed. The data set was then
    split into two sets where 80% of the samples were assigned to the
    training set and 20% to the testing set. This process was repeated
    for each of the 100 splits resulting in 100 training and testing
    datasets.
3.  VSEARCH Greengenes: Preprocessed data was clustered using VSEARCH
    v2.15.2 @rognes2016 directly to unprocessed Greengenes 97% OTU
    reference alignment consistent with how VSEARCH is typically used by
    the QIIME2 software for reference-based clustering @bolyen2019. The
    data was then subsampled to 10,000 reads and any novel OTUs from the
    test set were removed. The data set was then split into two sets
    where 80% of the samples were assigned to the training set and 20%
    to the testing set. This process was repeated for each of the 100
    splits resulting in 100 training and testing datasets.

@@

**Figure 2: Model perfomance of OptiFit self-reference workflow is as
good or better than other methods.** **A)** Area under the receiver
operating characteristic (AUROC) curve during cross-validation (train)
for the various workflows. **B)** AUROC on the test data for the various
workflows. The mean and standard deviation of the AUROC is represented
by the black dot and whiskers in panels A and B. The mean AUROC is
printed below the points. **C)** Averaged receiver operating
characteristic (ROC) curves. Lines represent the average true positive
rate for the range of false positive rates.

>>>>

1.  OptiFit Self: The preprocess data was split into the training and
    testing sets. The training set was clustered into OTUs using
    OptiClust, then the test set was fit to the OTUs of the training set
    using the OptiFit algorithm @sovacool2022. The OptiFit algorithm
    was run with method open so that any sequences that did not map to
    the existing OTU clusters would form new OTUs. The data was then
    subsampled to 10,000 reads and any novel OTUs from the test set were
    removed. This process was repeated for each of the 100 splits
    resulting in 100 training and testing datasets.
2.  OptiFit Greengenes: Reference sequences from the Greengenes database
    v13_8\_99 @desantis2006 were downloaded and processed with mothur by
    trimming to the V4 region and clustered *de novo* with OptiClust
    @westcott2017. The preprocessed data was fit to the clustered
    reference data using OptiFit with the method open to allow any
    sequences that did not map to the existing reference clusters would
    form new OTUs. The data was then subsampled to 10,000 reads and any
    novel OTUs from the test set were removed. The dataset was then
    split into two sets where 80% of the samples were assigned to the
    training set and 20% to the testing set. This process was repeated
    for each of the 100 splits resulting in 100 training and testing
    datasets.
3.  VSEARCH Greengenes: Preprocessed data was clustered using VSEARCH
    v2.15.2 @rognes2016 directly to unprocessed Greengenes 97% OTU
    reference alignment consistent with how VSEARCH is typically used by
    the QIIME2 software for reference-based clustering @bolyen2019. The
    data was then subsampled to 10,000 reads and any novel OTUs from the
    test set were removed. The dataset was then split into two sets
    where 80% of the samples were assigned to the training set and 20%
    to the testing set. This process was repeated for each of the 100
    splits resulting in 100 training and testing datasets.

@@

**Figure 2: Model performance of OptiFit self-reference workflow is as
good or better than other methods.** **A)** Area under the receiver
operating characteristic (AUROC) curve during cross-validation (train)
for the various workflows. **B)** AUROC on the test data for the various
workflows. The mean and standard deviation of the AUROC is represented
by the black dot and whiskers in panels A and B. The mean AUROC is
printed below the points. **C)** Averaged receiver operating
characteristic (ROC) curves. Lines represent the average true positive
rate for the range of false positive rates.

### 2023-04-13 submission/manuscript.tex
fix typos

  \begin{enumerate}
  \def\labelenumii{\arabic{enumii}.}
  \tightlist
  \item
    The preprocessed data was fit to the clustered reference data using
    OptiFit with the method open to allow any sequences that did not map
    to the existing reference clusters would form new OTUs. The data was
    then subsampled to 10,000 reads and any novel OTUs from the test set
    were removed. The data set was then split into two sets where 80\%
    of the samples were assigned to the training set and 20\% to the
    testing set. This process was repeated for each of the 100 splits
    resulting in 100 training and testing datasets.
  \end{enumerate}
\item
  VSEARCH Greengenes: Preprocessed data was clustered using VSEARCH
  v2.15.2 (\protect\hyperlink{ref-rognes2016}{4}) directly to
  unprocessed Greengenes 97\% OTU reference alignment consistent with
  how VSEARCH is typically used by the QIIME2 software for
  reference-based clustering (\protect\hyperlink{ref-bolyen2019}{5}).
  The data was then subsampled to 10,000 reads and any novel OTUs from
  the test set were removed. The data set was then split into two sets
  where 80\% of the samples were assigned to the training set and 20\%
  to the testing set. This process was repeated for each of the 100
  splits resulting in 100 training and testing datasets.
\end{enumerate}

@@

\textbf{Figure 2: Model perfomance of OptiFit self-reference workflow is
as good or better than other methods.} \textbf{A)} Area under the
receiver operating characteristic (AUROC) curve during cross-validation
(train) for the various workflows. \textbf{B)} AUROC on the test data
for the various workflows. The mean and standard deviation of the AUROC
is represented by the black dot and whiskers in panels A and B. The mean
AUROC is printed below the points. \textbf{C)} Averaged receiver
operating characteristic (ROC) curves. Lines represent the average true
positive rate for the range of false positive rates.

>>>>

  \begin{enumerate}
  \def\labelenumii{\arabic{enumii}.}
  \tightlist
  \item
    The preprocessed data was fit to the clustered reference data using
    OptiFit with the method open to allow any sequences that did not map
    to the existing reference clusters would form new OTUs. The data was
    then subsampled to 10,000 reads and any novel OTUs from the test set
    were removed. The dataset was then split into two sets where 80\% of
    the samples were assigned to the training set and 20\% to the
    testing set. This process was repeated for each of the 100 splits
    resulting in 100 training and testing datasets.
  \end{enumerate}
\item
  VSEARCH Greengenes: Preprocessed data was clustered using VSEARCH
  v2.15.2 (\protect\hyperlink{ref-rognes2016}{4}) directly to
  unprocessed Greengenes 97\% OTU reference alignment consistent with
  how VSEARCH is typically used by the QIIME2 software for
  reference-based clustering (\protect\hyperlink{ref-bolyen2019}{5}).
  The data was then subsampled to 10,000 reads and any novel OTUs from
  the test set were removed. The dataset was then split into two sets
  where 80\% of the samples were assigned to the training set and 20\%
  to the testing set. This process was repeated for each of the 100
  splits resulting in 100 training and testing datasets.
\end{enumerate}

@@

\textbf{Figure 2: Model performance of OptiFit self-reference workflow
is as good or better than other methods.} \textbf{A)} Area under the
receiver operating characteristic (AUROC) curve during cross-validation
(train) for the various workflows. \textbf{B)} AUROC on the test data
for the various workflows. The mean and standard deviation of the AUROC
is represented by the black dot and whiskers in panels A and B. The mean
AUROC is printed below the points. \textbf{C)} Averaged receiver
operating characteristic (ROC) curves. Lines represent the average true
positive rate for the range of false positive rates.
